How cryptic is cryptic diversity? Machine learning approaches to classifying morphological variation in \textit{Emys marmorata} (Testudinoidea, Emydidae).

### 2014-12-01 doc/draft_turtle.tex
filling out results

In classification studies, like this one, a common metric of performance is area under the receiver operating characteristic curve (AUC). AUC is an estimate of the relationship between the false positive and true positive rates, as opposed to just the true positive rate (accuracy). This relationship is especially useful in cases where misclassification needs to be minimized just as much as accurate classification, as in this study. AUC ranges between 0.5 and 1, with 0.5 indicating classification no better than random and 1 indicating perfect classification \citep{Hastie2009}.

@@

@@

% generalization
AUC values from model generalization, or estimating testing data set membership, does indicate a clear ``best'' classification scheme (Fig. \ref{fig:gen_hist}). While the scheme with two species has the greatest AUC point estimate for each modeling approach, this scheme is not significantly greater from any other except in some limited cases (e.g. LDA).

@@

\bibliographystyle{sysbio}
\bibliography{turtle,packages}

>>>>

In classification studies, like this one, a common metric of performance is area under the receiver operating characteristic curve (AUC). AUC is an estimate of the relationship between the false positive and true positive rates, as opposed to just the true positive rate (accuracy). This relationship is especially useful in cases where misclassification needs to be minimized just as much as accurate classification, as in this study. AUC ranges between 0.5 and 1, with 0.5 indicating classification no better than random and 1 indicating perfect classification \citep{Hastie2009}. AUC was used for both LDA and random forest models while multinomail logistic model selection was done using the likelihood based AICc \citep{Burnham2002a}.

@@

For both the LDA and multinomial logistic regression models, increasing model complexity increased cumulative percent variation necessary to best model the differing classification schemes FIGURE OF CUMMULATIVE. The high variance in optimal number of PCs used as predictors between schemes may represent either ease of fitting that particular model or that the data was too noising to accurately fit any model beyond the simplest.

As part of fitting a random forest model, a ranking of variable importance is also determined. Interestingly, the the order of variable importance is not the same as the order of principal components. This means that the principal components that best describe the differences between the various classes are not aligned with the principal components which describe the largest amount of variance. This result would be the case if variation between classes was extremely fine grained and not a part of the principal function or form of the plastron, which makes sense given that the plastron is involved in both protection and aquadynamics and not mate choice \citep{Germano2009,Holland1992,Lubcke2007,Rivera2008}.

The observed AUC values for all of the optimal models are very low, values near 0.5 indicate that a model is no better than completely random assigment TABLE OR FIGURE. This means that in very few cases are any of the proposed classification schemes are generally poor descriptors of the observed variation. It appears that the data set is overwhelmed by noise, making any accurate classifications difficult at best. This observation is cemented with the generalizations of the models to the testing data set.

@@

% generalization
AUC values from model generalization, or estimating testing data set membership, does not indicate a clear ``best'' classification scheme (Fig. \ref{fig:gen_hist}). While the scheme with two species has the greatest AUC point estimate for each modeling approach, this scheme is not significantly greater from any other except in some limited cases (e.g. LDA).

Mean AUC values for all model generalizations are, as expected, less than from the training set. This is expected because the low initial AUC values indicate little classification ability compared random, and the generalization to new data would only lead to loss in accuracy.

@@

Generalization and the unsupervised learning results are surprisingly congruous.

Plastron shape indicates no optimal classification scheme. A possible explanation for this that the genetic divergence associated with (sub)speciation is either not based on plastron morphology or local selective pressures due to hydrological regime overwhelming any possible morphological divergence.

\clearpage
\bibliographystyle{sysbio}
\bibliography{turtle,packages}
